### 2022-06-13 paper.tex
sc-BS-Seq -> scBS

@@

Here, we discuss bisulfite sequencing (sc-BS-Seq), a method to study DNA methylation. In brief, DNA is treated with bisulfite, which converts unmethylated cytosines to uracils which are read as thymine in subsequent PCR, while methylated cytosines are protected from conversion. After sequencing, these conversions allow to determine the methylation status of all cytosines covered by reads \citep{Frommer_1992}. Bisulfite sequencing can also be performed with single-cell resolution \citep{Smallwood_2014} and even in parallel with sc-RNA-Seq \citep{Clark2018}.

In the present paper, we discuss strategies to analyze sc-BS-Seq data, suggest improvement to current approaches, show their value in a benchmark, and present software to perform the improved analysis methods.

@@

To demonstrate the value of our proposed improvements, we obtained two previously published sc-BS-Seq data sets and re-analysed them.

@@

Since sc-BS-Seq protocols is still rather costly and labor-intensive, only few laboratories are currently able to obtain thousands of single cell methylomes.
To simulate a smaller data set, we subsampled the frontal cortex data set to 500 cells and repeated our analysis.
Here, our approach keeps its good score, while the simple approach falls behind (Figure \ref{figure:score}B).

@@

In brief, we store methylation of each chromosome in a matrix where each column represents a cell and each row represents a base pair. Methylated and unmethylated sites are coded as $+1$ and $-1$, respectively. Since the vast majority of values in this matrix are missing due to the sparsity of sc-BS-Seq data (and because rows for base-pairs not corresponding to a CpG site contain no data), we encode missing values as zero and then store the data in Compressed Sparse Row (CSR) format. This format does not explicitly store zeroes (here: missing values) and is optimized for row-wise access, which results in significant compression and allows fast access to the methylation status of genomic intervals.

@@

\vspace{1.6ex}
\noindent\hfil\rule{.6\columnwidth}{.2pt}\hfil

>>>>

\usepackage[compact]{titlesec}

@@

Here, we discuss bisulfite sequencing, a method to study DNA methylation. In brief, DNA is treated with bisulfite, which converts unmethylated cytosines to uracils which are read as thymine in subsequent PCR, while methylated cytosines are protected from conversion. After sequencing, these conversions allow to determine the methylation status of all cytosines covered by reads \citep{Frommer_1992}. Bisulfite sequencing can also be performed with single-cell resolution \citep{Smallwood_2014} and even in parallel with sc-RNA-Seq \citep{Clark2018}.

In the present paper, we discuss strategies to analyze scBS data, suggest improvement to current approaches, show their value in a benchmark, and present software to perform the improved analysis methods.

@@

To demonstrate the value of our proposed improvements, we obtained two previously published scBS data sets and re-analysed them.

@@

Since scBS protocols is still rather costly and labor-intensive, only few laboratories are currently able to obtain thousands of single cell methylomes.
To simulate a smaller data set, we subsampled the frontal cortex data set to 500 cells and repeated our analysis.
Here, our approach keeps its good score, while the simple approach falls behind (Figure \ref{figure:score}B).

@@

In brief, we store methylation of each chromosome in a matrix where each column represents a cell and each row represents a base pair. Methylated and unmethylated sites are coded as $+1$ and $-1$, respectively. Since the vast majority of values in this matrix are missing due to the sparsity of scBS data (and because rows for base-pairs not corresponding to a CpG site contain no data), we encode missing values as zero and then store the data in Compressed Sparse Row (CSR) format. This format does not explicitly store zeroes (here: missing values) and is optimized for row-wise access, which results in significant compression and allows fast access to the methylation status of genomic intervals.

@@

\vspace{1.4ex}
\noindent\hfil\rule{.6\columnwidth}{.2pt}\hfil
